---
title: "BT Baseline trials"
output: html_document
---

```{r setup, include=FALSE}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DISCLAIMER

This code was made by Carlos Rafael Gonzalez Soffner for the Final Thesis 
presented to the Bachelor of Data and Business Analytics. Anyone is welcome to
use this file or any part of it for their projects as long as they reference
this thesis, code or author.

Bests,

Carlos.


# LIBRARIES

The main libraries needed for modelling are readxl,BradleyTerry2 and dplyr.
Given that it is a classification task, it is likely that other libraries like 
pROC or ggplot2 might come at hand. Indeed, they were used in the initial
stages of the thesis to familiarise with the predictions, so they were left here
for future use.

```{r cars}
library(readxl)
library(BradleyTerry2)
library(openxlsx)
library(dplyr)
library(pROC)
library(caret)
library(tidyverse)
library(ggplot2)
```

# DATA

This chunk is used to read data and prepare the modelling. It includes:

  - list_STBT (list): list with training datasets.
  - list_ST_TEST (list): list with test datasets.
  - teams (vector): contains the teams in the datasets.
  - reference_wX (vector): contains the reference team for a window of X 
  seasons.


```{r}
list_STBT <- list(
  AFL2015 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2015.xlsx"),
  AFL2016 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2016.xlsx"),
  AFL2017 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2017.xlsx"),
  AFL2018 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2018.xlsx"),
  AFL2019 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2019.xlsx"),
  AFL2020 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2020.xlsx"),
  AFL2021 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2021.xlsx"),
  AFL2022 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2022.xlsx"),
  AFL2023 = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2023.xlsx"),
  AFLFULL = read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2024.xlsx")
)
list_ST_TEST <- list(
  AFL2015_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2015_TEST.xlsx"),
  AFL2016_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2016_TEST.xlsx"),
  AFL2017_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2017_TEST.xlsx"),
  AFL2018_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2018_TEST.xlsx"),
  AFL2019_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2019_TEST.xlsx"),
  AFL2020_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2020_TEST.xlsx"),
  AFL2021_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2021_TEST.xlsx"),
  AFL2022_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2022_TEST.xlsx"),
  AFL2023_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2023_TEST.xlsx"),
  AFLFULL_TEST = read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2024_TEST.xlsx")
)

teams <- c("teamADEL", "teamBL", "teamCARL", "teamCOLL", "teamESS", "teamFRE",
           "teamGCFC", "teamGEEL", "teamGWS", "teamHAW", "teamMELB", "teamNMFC",
           "teamPORT", "teamRICH", "teamSTK", "teamSYD", "teamWB", "teamWCE",
           "AT_HOME")  

reference_w1 <- c("GEEL", "PORT", "WB",
               "PORT", "PORT","GWS",
               "STK","STK","ADEL",
               "ADEL")
reference_w2 <- c("FRE", "STK", "HAW",
               "SYD", "ESS","SYD",
               "STK","STK","ADEL",
               "ADEL")
reference_w3 <- c("NMFC", "PORT", "HAW",
               "ESS", "STK","STK",
               "STK","ADEL","ADEL",
               "ADEL")
reference_w4 <- c("NMFC", "PORT", "ESS",
               "SYD", "GWS","FRE",
               "ADEL","ADEL","ADEL",
               "ADEL")

```

# FITTING STANDARD BT EXPANSIONS

The function fit_BTM_models is used to fit standard BT expansions to the
training datasets.

Parameters:
  - datasets (list): list with training datasets
  - teams (vector): contains the teams in the datasets.
  - ventana (numeric): size of the window to train the models
  - reference (vector): contains the reference team for a window of X 
  seasons.

Details:
  - Iterates over each dataset in the list and takes the correspondent reference
 team.
 - If the window is higher than 1 season, it takes the appropriate number of
 datasets, binds and formats them for modelling.
 - Sets the HOME_TEAM and AWAY_TEAM columns as factors and the HOME_WIN and 
 AWAY_WIN as numeric to avoid raising an Error.
 - Fits a BT model with BTm().
 - Iteratively builds two datasets: one with the season and AIC, and another
 with the estimated strength coefficients for each team. Contest-specific
 effects can be included but they will simply not be estimated and their rows
 will be left empty.
 
Returns:
  - list_models (list): list with fitted models so they are easier to access 
  when predicting
  - AIC_dataframe (data.frame): dataframe with the fitted models and their AIC.
  - standard_abilities (data.frame): dataframe with the estimated strengths for
  each season.

```{r}

fit_BTM_models <- function(datasets, teams, ventana, reference) {
  list_models <- list()
  # print('Creating dataframes')
  AIC_dataframe <- data.frame(
  model = character(),
  AIC = numeric(),
  predict_same = numeric(),
  predict_next = numeric(),
  stringsAsFactors = FALSE
  )
  
  standard_abilities <- data.frame(
    team = teams,
    stringsAsFactors = FALSE
  )
  reff = 0
  # print('Fitting models')
  # Fit model
  for (dataset_name in names(datasets)) {
    print(dataset_name)
    data_AFL <- datasets[[dataset_name]] 
    reff = reff + 1
    ########################
    #CHANGE: ventana effects
    ########################
    n_datasets <- min(ventana, length(datasets) - which(names(datasets) == dataset_name) + 1)
    if (n_datasets > 1) {
      for (i in 2:n_datasets) {
        next_dataset_name <- names(datasets)[which(names(datasets) == dataset_name) + (i - 1)]
        data_AFL <- rbind(data_AFL, datasets[[next_dataset_name]])
      }
      
      bound_datasets <- names(datasets)[
        which(names(datasets) %in% names(datasets)[which(
          names(datasets) == dataset_name):(
            which(names(datasets) == dataset_name) + n_datasets - 1)])]
      bound_datasets_string <- paste(bound_datasets, collapse = ", ")
      print(paste("Datasets binded:", bound_datasets_string))
      
      data_AFL <- data_AFL %>%
        group_by(HOME_TEAM, AWAY_TEAM) %>%
        summarise(
          HOME_WIN = sum(HOME_WIN),
          AWAY_WIN = sum(AWAY_WIN)) %>%
        ungroup()
    }  
    
    data_AFL$HOME_TEAM <- as.factor(data_AFL$HOME_TEAM)
    data_AFL$AWAY_TEAM <- as.factor(data_AFL$AWAY_TEAM)
    data_AFL$HOME_WIN <- as.numeric(as.character(data_AFL$HOME_WIN))
    data_AFL$AWAY_WIN <- as.numeric(as.character(data_AFL$AWAY_WIN))
    
    # print('Fitting ST')
    # print(reference[reff])
    standardBT <- BTm(cbind(HOME_WIN, AWAY_WIN), HOME_TEAM, AWAY_TEAM, 
                  data = data_AFL, id = "team", refcat = reference[reff]
                  )
    
    # print('summaryST')
    list_models[[dataset_name]] <- standardBT
    
    summary <- summary(standardBT)
    print(class(summary))
    AICC <- summary$aic
    # print(AICC)
    AIC_dataframe <- rbind(AIC_dataframe, data.frame(
      model = paste(dataset_name, "_model", sep = ""),
      AIC = AICC,
      predict_same = NA,
      predict_next = NA
    ))
    
    coef_names <- rownames(summary$coefficients)
    coef_values <- summary$coefficients[, 1]
    coef_pvalues <- summary$coefficients[, 4]
    standard_abilities[[paste("ability_", dataset_name, 
                              sep = "")]] <- numeric(nrow(standard_abilities))
    standard_abilities[[paste("pvalue_", dataset_name, 
                              sep = "")]] <- numeric(nrow(standard_abilities))
    for (i in seq_along(coef_names)) {
      coef_name <- coef_names[i]
      coef_value <- coef_values[i]
      coef_pvalue <- coef_pvalues[i]
      if (coef_name %in% teams) {
        standard_abilities[standard_abilities$team == coef_name, 
                           paste("ability_", dataset_name, sep = "")] <- coef_value
        standard_abilities[standard_abilities$team == coef_name, 
                           paste("pvalue_", dataset_name, sep = "")] <- coef_pvalue
      }
    }
  }
  
  result <- list(
    list_models = list_models,
    AIC_dataframe = AIC_dataframe,
    standard_abilities = standard_abilities
  )
  return(result)
  
}
```


```{r}
resu <- fit_BTM_models(list_STBT, teams, 4, reference_w4)
resu
```

```{r}
write.xlsx(resu$AIC_dataframe, file = "E:/TFG DATA/RESULTS/AIC_PREDS_ST_w4.xlsx", 
           row.names = FALSE)
write.xlsx(resu$standard_abilities, file = "E:/TFG DATA/RESULTS/COEFS_ST_w4.xlsx", 
           row.names = FALSE)
```

# PREDICTING WITH STANDARD BT EXPANSIONS

The function d_binder_test is used to bind two or more test datasets to test
the model on a window of more than one season.

predict_same_next_ST is used to make predictions with the previously fitted
models.

Parameters:
  - model_ST (model): Bradley-Terry model.
  - same_data (data.frame): training data.
  - next_data (data.frame): test data.

Details:
  - Predictions are done at the 'response' level. It predicts the probability
  between 0 and 1 of the HOME_TEAM winning. The HOME_TEAM is predicted to win
  if the probability is > 0.5.
 
Returns:
  - accuracy_now (numeric): classification accuracy over train set
  - accuracy_next (numeric): classification accuracy over test set
  - conf_matrix_now: confusion matrix for the predictions over the train set.
  - conf_matrix_next: confusion matrix for the predictions over the test set.


```{r}
d_binder_test <- function(data_list){
  data_AFL <- do.call(rbind, data_list)
  
  #data_AFL <- data_AFL %>%
  #group_by(HOME_TEAM, AWAY_TEAM) %>%
  #summarise(
  #  HOME_WIN = sum(HOME_WIN),
  #  AWAY_WIN = sum(AWAY_WIN)) %>%
  #ungroup()

  data_AFL$HOME_TEAM <- as.factor(data_AFL$HOME_TEAM)
  data_AFL$AWAY_TEAM <- as.factor(data_AFL$AWAY_TEAM)
  data_AFL$HOME_WIN <- as.numeric(as.character(data_AFL$HOME_WIN))
  data_AFL$AWAY_WIN <- as.numeric(as.character(data_AFL$AWAY_WIN))
  
  return(data_AFL)
  
}


predict_same_next_ST <- function(model_ST, same_data,next_data) {
  
  same_data$HOME_TEAM <- as.factor(same_data$HOME_TEAM)
  same_data$AWAY_TEAM <- as.factor(same_data$AWAY_TEAM)
  same_data$HOME_WIN <- as.numeric(as.character(same_data$HOME_WIN))
  same_data$AWAY_WIN <- as.numeric(as.character(same_data$AWAY_WIN))
  next_data$HOME_TEAM <- as.factor(next_data$HOME_TEAM)
  next_data$AWAY_TEAM <- as.factor(next_data$AWAY_TEAM)
  next_data$HOME_WIN <- as.numeric(as.character(next_data$HOME_WIN))
  next_data$AWAY_WIN <- as.numeric(as.character(next_data$AWAY_WIN))
  
  
  pred_now_p <- predict(model_ST, se.fit = TRUE, level = 0, 
                      type = "response", newdata = same_data)$fit
  pred_now_b <- ifelse(pred_now_p > 0.5, 1, 0)
  pred_next_p <- predict(model_ST, se.fit = TRUE, level = 0, 
                      type = "response", newdata = next_data)$fit
  pred_next_b <- ifelse(pred_next_p > 0.5, 1, 0)
  
  accuracy_now <- mean(pred_now_b == same_data$HOME_WIN)
  accuracy_next <- mean(pred_next_b == next_data$HOME_WIN)
  conf_matrix_now <- confusionMatrix(data = factor(pred_now_b,levels = c(0, 1)), 
                               reference = factor(same_data$HOME_WIN,levels = c(0, 1)),
                               positive = '1')
  conf_matrix_next <- confusionMatrix(data = factor(pred_next_b,levels = c(0, 1)), 
                             reference = factor(next_data$HOME_WIN,levels = c(0, 1)),
                             positive = '1')
  
  result <- list(
    accuracy_now = accuracy_now,
    accuracy_next = accuracy_next,
    conf_matrix_now = conf_matrix_now,
    conf_matrix_next = conf_matrix_next
  )
  return(result)
  
}
```


```{r}

bindi<-d_binder_test(list(list_ST_TEST$AFL2023_TEST,
                          list_ST_TEST$AFL2020_TEST,
                          list_ST_TEST$AFL2021_TEST,
                         list_ST_TEST$AFL2022_TEST))

preds_ST<-predict_same_next_ST(resu$list_models$AFL2020,
                               bindi,
                               list_ST_TEST$AFL2023_TEST)
preds_ST$conf_matrix_now
preds_ST$conf_matrix_next
```

# FITTING CONTEST-SPECIFIC BT EXPANSIONS

The function fit_BTM_CS is used to fit contest-specufic BT expansions to the
training datasets. It works just like fit_BTM_models, with the difference being
the reprocessing of the HOME_TEAM and AWAY_TEAM variables as dataframes with
the columns 'team' and AT_HOME (1 for HOME_TEAM, 0 for AWAY_TEAM).


```{r}
fit_BTM_CS <- function(datasets, teams, ventana, reference) {
  list_models <- list()
  # print('Creating dataframes')
  AIC_dataframe <- data.frame(
  model = character(),
  AIC = numeric(),
  predict_same = numeric(),
  predict_next = numeric(),
  stringsAsFactors = FALSE
  )
  
  standard_abilities <- data.frame(
    team = teams,
    stringsAsFactors = FALSE
  )
  reff = 0
  # print('Fitting models')
  # Fit model
  for (dataset_name in names(datasets)) {
    reff = reff + 1
    print(dataset_name)
    data_AFL <- datasets[[dataset_name]] 
    ########################
    #CHANGE: ventana effects
    ########################
    n_datasets <- min(ventana, length(datasets) - which(names(datasets) == dataset_name) + 1)
    if (n_datasets > 1) {
      for (i in 2:n_datasets) {
        next_dataset_name <- names(datasets)[which(names(datasets) == dataset_name) + (i - 1)]
        data_AFL <- rbind(data_AFL, datasets[[next_dataset_name]])
      }
      bound_datasets <- names(datasets)[
        which(names(datasets) %in% names(datasets)[which(
          names(datasets) == dataset_name):(
            which(names(datasets) == dataset_name) + n_datasets - 1)])]
      bound_datasets_string <- paste(bound_datasets, collapse = ", ")
      print(paste("Datasets binded:", bound_datasets_string))
      data_AFL <- data_AFL %>%
        group_by(HOME_TEAM, AWAY_TEAM) %>%
        summarise(
          HOME_WIN = sum(HOME_WIN),
          AWAY_WIN = sum(AWAY_WIN)) %>%
        ungroup()
    }
    
    data_AFL$HOME_TEAM <- as.factor(data_AFL$HOME_TEAM)
    data_AFL$AWAY_TEAM <- as.factor(data_AFL$AWAY_TEAM)
    data_AFL$HOME_WIN <- as.numeric(as.character(data_AFL$HOME_WIN))
    data_AFL$AWAY_WIN <- as.numeric(as.character(data_AFL$AWAY_WIN))
    data_AFL$HOME_TEAM <- data.frame(team = data_AFL$HOME_TEAM,
                                     AT_HOME = 1)
    data_AFL$AWAY_TEAM <- data.frame(team = data_AFL$AWAY_TEAM
                                      , AT_HOME = 0)
    
    # print('Fitting ST-CS')
    standardBT_HOME<- BTm(outcome = cbind(HOME_WIN, AWAY_WIN), 
                          player1 = HOME_TEAM, player2 = AWAY_TEAM, 
                          formula = ~team + AT_HOME, id = "team",
                           refcat = reference[reff],
                      data = data_AFL)
    
    # print('summaryST')
    list_models[[dataset_name]] <- standardBT_HOME
    
    summary <- summary(standardBT_HOME)
    # print(summary)
    AICC <- summary$aic
    # print(AICC)
    AIC_dataframe <- rbind(AIC_dataframe, data.frame(
      model = paste(dataset_name, "_model", sep = ""),
      AIC = AICC,
      predict_same = NA,
      predict_next = NA
    ))
    
    coef_names <- rownames(summary$coefficients)
    coef_values <- summary$coefficients[, 1]
    coef_pvalues <- summary$coefficients[, 4]
    standard_abilities[[paste("ability_", dataset_name, 
                              sep = "")]] <- numeric(nrow(standard_abilities))
    standard_abilities[[paste("pvalue_", dataset_name, 
                              sep = "")]] <- numeric(nrow(standard_abilities))
    for (i in seq_along(coef_names)) {
      coef_name <- coef_names[i]
      coef_value <- coef_values[i]
      coef_pvalue <- coef_pvalues[i]
      if (coef_name %in% teams) {
        standard_abilities[standard_abilities$team == coef_name, 
                           paste("ability_", dataset_name, sep = "")] <- coef_value
        standard_abilities[standard_abilities$team == coef_name, 
                           paste("pvalue_", dataset_name, sep = "")] <- coef_pvalue
      }
    }
  }
  
  result <- list(
    list_models = list_models,
    AIC_dataframe = AIC_dataframe,
    standard_abilities = standard_abilities
  )
  return(result)
  
}
```


```{r}
resu_cs <- fit_BTM_CS(list_STBT, teams, 4, reference = reference_w4)
# resu_cs
```
```{r}
resu_cs$standard_abilities
```


```{r}
write.xlsx(resu_cs$AIC_dataframe, 
           file = "E:/TFG DATA/RESULTS/AIC_PREDS_CS_w4.xlsx", row.names = FALSE)
write.xlsx(resu_cs$standard_abilities, 
           file = "E:/TFG DATA/RESULTS/COEFS_CS_w4.xlsx", row.names = FALSE)
```

# PREDICTING CONTEST-SPECIFIC BT EXPANSIONS

The function predict_same_next_CS is used to make predictions with the 
previously fitted models. It works just like predict_same_next_ST, with the 
difference being the reprocessing of the HOME_TEAM and AWAY_TEAM variables as 
dataframes with the columns 'team' and AT_HOME (1 for HOME_TEAM, 0 for 
AWAY_TEAM).

```{r}
predict_same_next_CS <- function(model_CS, same_data,next_data) {
  
  same_data$HOME_TEAM <- as.factor(same_data$HOME_TEAM)
  same_data$AWAY_TEAM <- as.factor(same_data$AWAY_TEAM)
  same_data$HOME_WIN <- as.numeric(as.character(same_data$HOME_WIN))
  same_data$AWAY_WIN <- as.numeric(as.character(same_data$AWAY_WIN))
  same_data$HOME_TEAM <- data.frame(team = same_data$HOME_TEAM,
                                     AT_HOME = 1)
  same_data$AWAY_TEAM <- data.frame(team = same_data$AWAY_TEAM
                                      , AT_HOME = 0)
  next_data$HOME_TEAM <- as.factor(next_data$HOME_TEAM)
  next_data$AWAY_TEAM <- as.factor(next_data$AWAY_TEAM)
  next_data$HOME_WIN <- as.numeric(as.character(next_data$HOME_WIN))
  next_data$AWAY_WIN <- as.numeric(as.character(next_data$AWAY_WIN))
  next_data$HOME_TEAM <- data.frame(team = next_data$HOME_TEAM,
                                     AT_HOME = 1)
  next_data$AWAY_TEAM <- data.frame(team = next_data$AWAY_TEAM
                                      , AT_HOME = 0)
  
  
  pred_now_p <- predict(model_CS, se.fit = TRUE, level = 0, 
                      type = "response", newdata = same_data)$fit
  pred_now_b <- ifelse(pred_now_p > 0.5, 1, 0)
  pred_next_p <- predict(model_CS, se.fit = TRUE, level = 0, 
                      type = "response", newdata = next_data)$fit
  pred_next_b <- ifelse(pred_next_p > 0.5, 1, 0)
  
  accuracy_now <- mean(pred_now_b == same_data$HOME_WIN)
  accuracy_next <- mean(pred_next_b == next_data$HOME_WIN)
  conf_matrix_now <- confusionMatrix(data = factor(pred_now_b,levels = c(0, 1)), 
                               reference = factor(same_data$HOME_WIN,levels = c(0, 1)),
                               positive = '1')
  conf_matrix_next <- confusionMatrix(data = factor(pred_next_b,levels = c(0, 1)), 
                             reference = factor(next_data$HOME_WIN,levels = c(0, 1)),
                             positive = '1')
  
  result <- list(
    accuracy_now = accuracy_now,
    accuracy_next = accuracy_next,
    conf_matrix_now = conf_matrix_now,
    conf_matrix_next = conf_matrix_next
  )
  return(result)
  
}
```


```{r}
test_list <- list(list_ST_TEST$AFL2023_TEST,
                  list_ST_TEST$AFL2020_TEST,
                  list_ST_TEST$AFL2021_TEST,
                  list_ST_TEST$AFL2022_TEST)
bindi <- do.call(rbind, test_list)
preds_CS <- predict_same_next_CS(resu_cs$list_models$AFL2019,
                                 bindi, 
                                 list_ST_TEST$AFL2023_TEST)
preds_CS$conf_matrix_now
preds_CS$conf_matrix_next
```
