---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DISCLAIMER

This code was made by Carlos Rafael Gonzalez Soffner for the Final Thesis 
presented to the Bachelor of Data and Business Analytics. Anyone is welcome to
use this file or any part of it for their projects as long as they reference
this thesis, code or author.

Bests,

Carlos.

# LIBRARIES

The AFL data used in this thesis is fetched through the fitzRoy package.
Ensure it is installed and imported before running the cells below.

```{r}
library(fitzRoy)
```

# FETCHING RESULTS

Results are fetched from AFLtables with the function fetch_results(). A for 
loop is used to retrieve results from 2012 to 2023, even if the final seasons
used in this thesis are 2015-23. Each season is retrieved individually and 
appended into a list. All datasets in the list are binded with rbind.

```{r}
results_list <- list()
for (season in 2012:2023) {
  print(paste("Fetching results for", season))
  results <- fetch_results(season, comp = "AFLM", 
                           round_number = NULL,
                           source = "afltables")
  results_list[[as.character(season)]] <- results
  print(paste("Fetched results for", season))
}
results_afl_tables_12TO23 <- do.call(rbind, results_list)

```

```{r}
write.csv(results_afl_tables_12TO23, 
          file = "E:/TFG DATA/results_afl_tables_12TO23.csv", 
          row.names = FALSE)
```


```{r}
##########################################################################
# IGNORE - TRIALS
# These lines are commented but feel free to use them to visualise the data
# from one seasons and have an idea of what is being fetched without having
# to wait for the for loop.
##########################################################################

# results_season_23_afltables <- fetch_results(2023, comp = "AFLM", 
#                                             round_number = NULL,
#                                             source = "afltables")
# 
# results_season_22_afltables <- fetch_results(2022, comp = "AFLM", 
#                                             round_number = NULL,
#                                             source = "afltables")
# 
# results_season_21_afltables <- fetch_results(2021, comp = "AFLM", 
#                                             round_number = NULL,
#                                             source = "afltables")
# 
# results_afl_tables_21TO23 <- rbind(results_season_21_afltables, 
#                                    results_season_22_afltables, 
#                                    results_season_23_afltables)
# 
# View(results_afl_tables_21TO23)
```


```{r}
# write.csv(results_afl_tables_21TO23, 
#           file = "E:/TFG DATA/results_afl_tables_21TO23.csv", 
#           row.names = FALSE)
```


# FETCHING LADDER

Ladder data for seasons 2015 to 2023 is fetched form the AFL with the function
fetch_ladder(). Each season is fetched individually since the number of columns
might change season-to-season or even in the middle of the season, probably as 
a consequence from new rules or standards to document data in the AFL. The file
with the ladder data for all seasons is created by running rbind on the
individually fetched seasons.


```{r}


ladder23 <- data.frame()
season <-2023
for (round in 1:24) {
  print(paste("Fetching", round, nrow(ladder23), ncol(ladder23)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder23)>0){
    missing_columns <- setdiff(names(ladder23), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder23))
    for (col in missing_columns) {
      ladder23[[col]] <- NA
    }
    
  }
  
  ladder23  <- rbind(ladder23 , round_ladder)
}

View(ladder23)

```

```{r}
ladder22 <- data.frame()
season <-2022
for (round in 1:23) {
  print(paste("Fetching", round, nrow(ladder22), ncol(ladder22)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder22)>0){
    missing_columns <- setdiff(names(ladder22), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder22))
    for (col in missing_columns) {
      ladder22[[col]] <- NA
    }
    
  }
  
  ladder22  <- rbind(ladder22 , round_ladder)
}

View(ladder22)
```


```{r}
ladder21 <- data.frame()
season <-2021
for (round in 1:23) {
  print(paste("Fetching", round, nrow(ladder21), ncol(ladder21)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder21)>0){
    missing_columns <- setdiff(names(ladder21), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder21))
    for (col in missing_columns) {
      ladder21[[col]] <- NA
    }
    
  }
  
  ladder21  <- rbind(ladder21 , round_ladder)
}

View(ladder21)
```


```{r}
ladder20 <- data.frame()
season <-2020
for (round in 1:18) {
  print(paste("Fetching", round, nrow(ladder20), ncol(ladder20)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder20)>0){
    missing_columns <- setdiff(names(ladder20), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder20))
    for (col in missing_columns) {
      ladder20[[col]] <- NA
    }
    
  }
  
  ladder20  <- rbind(ladder20 , round_ladder)
}

View(ladder20)
```
```{r}
ladder19 <- data.frame()
season <-2019
for (round in 1:23) {
  print(paste("Fetching", round, nrow(ladder19), ncol(ladder19)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder19)>0){
    missing_columns <- setdiff(names(ladder19), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder19))
    for (col in missing_columns) {
      ladder19[[col]] <- NA
    }
    
  }
  
  ladder19  <- rbind(ladder19 , round_ladder)
}

View(ladder19)
```

```{r}
ladder18 <- data.frame()
season <-2018
for (round in 1:23) {
  print(paste("Fetching", round, nrow(ladder18), ncol(ladder18)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder18)>0){
    missing_columns <- setdiff(names(ladder18), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder18))
    for (col in missing_columns) {
      ladder18[[col]] <- NA
    }
    
  }
  
  ladder18  <- rbind(ladder18 , round_ladder)
}

View(ladder18)
```

```{r}
ladder17 <- data.frame()
season <-2017
for (round in 1:23) {
  print(paste("Fetching", round, nrow(ladder17), ncol(ladder17)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder17)>0){
    missing_columns <- setdiff(names(ladder17), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder17))
    for (col in missing_columns) {
      ladder17[[col]] <- NA
    }
    
  }
  
  ladder17  <- rbind(ladder17 , round_ladder)
}

View(ladder17)
```


```{r}
ladder16 <- data.frame()
season <-2016
for (round in 1:23) {
  print(paste("Fetching", round, nrow(ladder16), ncol(ladder16)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder16)>0){
    missing_columns <- setdiff(names(ladder16), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder16))
    for (col in missing_columns) {
      ladder16[[col]] <- NA
    }
    
  }
  
  ladder16  <- rbind(ladder16 , round_ladder)
}

View(ladder16)
```

```{r}
ladder15 <- data.frame()
season <-2015
for (round in 1:23) {
  print(paste("Fetching", round, nrow(ladder15), ncol(ladder15)))
  round_ladder <- fetch_ladder(season, comp = "AFLM", round_number = round, source = "AFL")
  
  if (nrow(ladder15)>0){
    missing_columns <- setdiff(names(ladder15), names(round_ladder))
    for (col in missing_columns) {
      round_ladder[[col]] <- NA
    }
    missing_columns <- setdiff(names(round_ladder),names(ladder15))
    for (col in missing_columns) {
      ladder15[[col]] <- NA
    }
    
  }
  
  ladder15  <- rbind(ladder15 , round_ladder)
}

View(ladder15)
```

```{r}
missing_columns <- setdiff(names(ladder22), names(ladder20))
for (col in missing_columns) {
  print(col)
  ladder21[[col]] <- NA
  ladder20[[col]] <- NA
}  


```
```{r}
missing_columns <- setdiff(names(ladder20), names(ladder19))
for (col in missing_columns) {
  print(col)
    ladder19[[col]] <- NA
    ladder18[[col]] <- NA
    ladder17[[col]] <- NA
    ladder16[[col]] <- NA
    ladder15[[col]] <- NA
}  
```

```{r}
LADDER_15TO23 <- rbind(ladder15, ladder16, ladder17,
                       ladder18, ladder19, ladder20,
                       ladder21, ladder22, ladder23)
```


```{r}
write.csv(LADDER_15TO23, file = "E:/TFG DATA/LADDER_15TO23.csv", 
          row.names = FALSE)
```

# FETCHING PLAYER DATA

Player data (performance indicators) is fetched from the AFL with the 
fetch_player_stats_afl() function.
Each season is fetched individually so it is easier to add a SEASON column to 
identify the each season when binding the data. They are finally concatenated
together. The middle cell includes a code to compare the columns in each season
in case they are different.

```{r}
player15 <- fetch_player_stats_afl(season = 2015
                                      , comp = "AFLM")
player16 <- fetch_player_stats_afl(season = 2016
                                      , comp = "AFLM")
player17 <- fetch_player_stats_afl(season = 2017
                                      , comp = "AFLM")

player18 <- fetch_player_stats_afl(season = 2018
                                      , comp = "AFLM")
player19 <- fetch_player_stats_afl(season = 2019
                                      , comp = "AFLM")
player20 <- fetch_player_stats_afl(season = 2020
                                      , comp = "AFLM")

player21 <- fetch_player_stats_afl(season = 2021
                                      , comp = "AFLM")
player22 <- fetch_player_stats_afl(season = 2022
                                      , comp = "AFLM")
player23 <- fetch_player_stats_afl(season = 2023
                                      , comp = "AFLM")
player15$SEASON <- 2015
player16$SEASON <- 2016
player17$SEASON <- 2017
player18$SEASON <- 2018
player19$SEASON <- 2019
player20$SEASON <- 2020
player21$SEASON <- 2021
player22$SEASON <- 2022
player23$SEASON <- 2023

```

```{r}
missing_columns <- setdiff(names(player16), names(player17))
for (col in missing_columns) {
  print(col)
  player17[[col]] <- NA
  player18[[col]] <- NA
  player19[[col]] <- NA
  player20[[col]] <- NA
  player21[[col]] <- NA
  player22[[col]] <- NA
  player23[[col]] <- NA
}  
```


```{r}

PIS_21TO23 <- rbind(player15,
                    player16,
                    player17,
                    player18,
                    player19,
                    player20,
                    player21,
                    player22,
                    player23)

```

```{r}
 write.csv(PIS_21TO23, file = "E:/TFG DATA/PIS_15TO23.csv", 
          row.names = FALSE)
```

