---
title: "BT_TS_TV"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DISCLAIMER

This code was made by Carlos Rafael Gonzalez Soffner for the Final Thesis 
presented to the Bachelor of Data and Business Analytics. Anyone is welcome to
use this file or any part of it for their projects as long as they reference
this thesis, code or author.

Bests,

Carlos.


# LIBRARIES

The main libraries needed for modelling are readxl,BradleyTerry2 and dplyr.
Given that it is a classification task, it is likely that other libraries like 
pROC or ggplot2 might come at hand. Indeed, they were used in the initial
stages of the thesis to familiarise with the predictions, so they were left here
for future use.

```{r cars}
library(readxl)
library(BradleyTerry2)
library(openxlsx)
library(dplyr)
library(pROC)
library(caret)
library(tidyverse)
library(ggplot2)
```

# DATA PREPARATION

data_build prepares the data for modelling complex BT expansions.

Parameters:
  - homepath (string): path to home team data
  - awaypath (string): path to away team data
  - fixedpath (string): path to fixed effects data. For this thesis, none of
  this features are used. It should have at least a column with the team names.
  
Returns:
 - afldata (list): list with home, away and fixed dataframe

```{r}
data_build <- function(homepath, awaypath, fixedpath) {
  
  home <- as.data.frame(read_excel(homepath))
  away <- as.data.frame(read_excel(awaypath))
  fixed <- as.data.frame(read_excel(fixedpath))

  row.names(fixed) <- fixed$TEAM
  home$TEAM <- as.factor(home$TEAM)
  away$TEAM <- as.factor(away$TEAM)
  
  afldata <- list(home = home, away = away, fixed = fixed)
  
  return(afldata) 
}
```

# FITTING TEAM-SPECIFIC, TIME-VARIANT BT EXPANSION

fit_BTM_TSTV is used to fit the team-specific, time-variant BT expansion to a
particular season or window of seasons.

Parameters:
  - afldata (list): list with home, away and fixed dataframe
  - feats (vector): vector of features to be modelled.

Details:
  - Trains a BT model on the train data for each feature. If the feature is 
  found to be significant at the 5% level, it appends it to a list of 
  significant features.
  - Then, it trains a model on all the significant features and takes that with
  the highest p-value. If it is not below 0.05, if discards the feature and
  retrains the model. Otherwise, it finishes.
  - If all features are discarded it returns an empty list. This will be useful
  when making round-by-round predictions.

Returns:
 - model (model): final model with significant features.
 - summary (summary object): summary of the model returned.

```{r}
fit_BTM_TSTV <- function(afldata, feats) {
  
  # Init
  sig_feats <- list()
  all_mod <- 0
  # Fit single-feature models
  for (feature in feats) {
    # print(feature)
    # Fit model
    btm_model <- BTm(outcome = WIN,
                     player1 = home,
                     player2 = away,
                     formula = as.formula(paste("~", feature, "+ (1|TEAM)")),
                     id = "TEAM",
                     data = afldata)
    # P-value and significance, add feature if significant
    summary_coef <- as.data.frame(as.array(summary(btm_model))[6])
    av <- nchar(toString(as.array(summary(btm_model))[6]))
    if (av < 10) {
      summary_coef <- as.data.frame(as.array(summary(btm_model))[11])
      }
    # print(summary_coef)
    colnames(summary_coef) <- c("estimate", "stderror", "zvalue", "pvalue")
    p_value <- summary_coef$pvalue
    if (p_value < 0.05) {
      # cat("Feature", feature, "is significant with coefficient", 
      #     coef(btm_model)[feature], "and p-value", p_value, "\n")
      sig_feats <- c(sig_feats, feature)
    }
  }
  

  if (length(sig_feats) > 0) {
    
      while (all_mod == 0) {
        formula_all <- as.formula(paste("~", paste(sig_feats, collapse = " + "),
                                    "+ (1|TEAM)"))
        btm_model_all <- BTm(outcome = WIN,
                         player1 = home,
                         player2 = away,
                         formula = formula_all,
                         id = "TEAM",
                         data = afldata)
        
        summary_all <- as.data.frame(as.array(summary(btm_model_all))[6])
        av <- nchar(toString(as.array(summary(btm_model_all))[6]))
        if (av < 10) {
          summary_all <- as.data.frame(as.array(summary(btm_model_all))[11])
          }
        # print(summary_all)
        colnames(summary_all) <- c("estimate", "stderror", "zvalue", "pvalue")
        row_max_pvalue <- which.max(summary_all$pvalue)
        max_pvalue <- summary_all$pvalue[row_max_pvalue]
        if (max_pvalue > 0.05) {
          # cat("Feature", rownames(summary_all)[row_max_pvalue],
          #     "has the highest p-value:", max_pvalue, "\n")
          feature_to_remove <- rownames(summary_all)[row_max_pvalue]
          sig_feats <- sig_feats[sig_feats != feature_to_remove]
        } else {
            all_mod <- 1
          }
    
      }
  } else {
      btm_model_all <- list()
    }
  
  return(list(model = btm_model_all, 
              summary = summary(btm_model_all)))
}
```

# EXAMPLES

```{r}
homepath <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_FULL.xlsx"
awaypath <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_FULL.xlsx"
fixedpath <- "E:/TFG DATA/last_ladder_position.xlsx"
feats_ALL <- list('CONSECUTIVE_WINS', 'CONSECUTIVE_LOSSES', 'LG_WON',
       'LADDER_POSITION_DIFF', 'LADDERLY_POSITION_DIFF', 'POINTSAGAINST_DIFF',
       'CLANGERS_CSUM_DIFF', 'FREES_AGAINST_CSUM_DIFF', 'TURNOVERS_CSUM_DIFF',
       'CONTEST_DEFENSIVE_LOSS_CSUM_DIFF', 'WINS_CUMMULATIVE_DIFF',
       'PERCENTAGE_DIFF', 'POINTSFOR_DIFF', 'DISPOSALS_CSUM_DIFF',
       'DISPOSALS_EFFECTIVE_CSUM_DIFF', 'KICKS_CSUM_DIFF',
       'KICKS_EFFECTIVE_CSUM_DIFF', 'HANDBALLS_CSUM_DIFF', 'MARKS_CSUM_DIFF',
       'MARKS_INSIDE50_CSUM_DIFF', 'MARKS_CONTESTED_CSUM_DIFF',
       'MARKS_ONLEAD_CSUM_DIFF', 'MARKS_INTERCEPT_CSUM_DIFF',
       'BOUNCES_CSUM_DIFF', 'METRES_GAINED_CSUM_DIFF', 'POSESSIONS_CSUM_DIFF',
       'POSESSIONS_CONTESTED_CSUM_DIFF', 'POSESSIONS_UNCONTESTED_CSUM_DIFF',
       'GETS_GROUNDBALL_CSUM_DIFF', 'GETS_GROUNDBALL50_CSUM_DIFF',
       'INSIDE50_CSUM_DIFF', 'GOALS_SHOTS_CSUM_DIFF',
       'SCORE_LAUNCHES_CSUM_DIFF', 'CLEARANCES_CSUM_DIFF',
       'CLEARANCES_CENTRE_CSUM_DIFF', 'CLEARANCES_STOPPAGE_CSUM_DIFF',
       'HITOUTS_ADVANTAGE_CSUM_DIFF', 'HITOUTS_WIN_RATE_CSUM_DIFF',
       'CONTEST_OFFENSIVE_WIN_CSUM_DIFF', 'TACKLES_CSUM_DIFF',
       'TACKLES_INSIDE50_CSUM_DIFF', 'ONE_PERCENTERS_CSUM_DIFF',
       'REBOUND_INSIDE50S_CSUM_DIFF', 'INTERCEPTS_CSUM_DIFF',
       'PRESSURE_CSUM_DIFF', 'PRESSURE_DEFENSEHALF_CSUM_DIFF',
       'SPOILS_CSUM_DIFF', 'DISPOSALS_EFFICIENCY_CSUM_DIFF',
       'KICKS_EFFICIENCY_CSUM_DIFF', 'KICK2HANDBALL_CSUM_DIFF',
       'POSESSIONS_CONTESTED_RATE_CSUM_DIFF', 'GOALS_ACCURACY_CSUM_DIFF',
       'HITOUTS_ADVANTAGE_RATE_CSUM_DIFF',
       'CONTEST_OFFENSIVE_WIN_RATE_CSUM_DIFF',
       'CONTEST_DEFENSIVE_LOSS_RATE_CSUM_DIFF', 'AT_HOME', 'HOMEGROUND',
       'INTERSTATE')
feats_L4 <- list('CONSECUTIVE_WINS', 'CONSECUTIVE_LOSSES', 'LG_WON', 'L4G_WINS',
       'LADDER_POSITION_DIFF', 'LADDERLY_POSITION_DIFF', 'POINTSAGAINST_DIFF',
       'CLANGERS_L4_CSUM_DIFF', 'FREES_AGAINST_L4_CSUM_DIFF',
       'TURNOVERS_L4_CSUM_DIFF', 'CONTEST_DEFENSIVE_LOSS_L4_CSUM_DIFF',
       'PERCENTAGE_DIFF', 'POINTSFOR_DIFF', 'DISPOSALS_L4_CSUM_DIFF',
       'DISPOSALS_EFFECTIVE_L4_CSUM_DIFF', 'KICKS_L4_CSUM_DIFF',
       'KICKS_EFFECTIVE_L4_CSUM_DIFF', 'HANDBALLS_L4_CSUM_DIFF',
       'MARKS_L4_CSUM_DIFF', 'MARKS_INSIDE50_L4_CSUM_DIFF',
       'MARKS_CONTESTED_L4_CSUM_DIFF', 'MARKS_ONLEAD_L4_CSUM_DIFF',
       'MARKS_INTERCEPT_L4_CSUM_DIFF', 'BOUNCES_L4_CSUM_DIFF',
       'METRES_GAINED_L4_CSUM_DIFF', 'POSESSIONS_L4_CSUM_DIFF',
       'POSESSIONS_CONTESTED_L4_CSUM_DIFF',
       'POSESSIONS_UNCONTESTED_L4_CSUM_DIFF', 'GETS_GROUNDBALL_L4_CSUM_DIFF',
       'GETS_GROUNDBALL50_L4_CSUM_DIFF', 'INSIDE50_L4_CSUM_DIFF',
       'GOALS_SHOTS_L4_CSUM_DIFF', 'SCORE_LAUNCHES_L4_CSUM_DIFF',
       'CLEARANCES_L4_CSUM_DIFF', 'CLEARANCES_CENTRE_L4_CSUM_DIFF',
       'CLEARANCES_STOPPAGE_L4_CSUM_DIFF', 'HITOUTS_ADVANTAGE_L4_CSUM_DIFF',
       'HITOUTS_WIN_RATE_L4_CSUM_DIFF', 'CONTEST_OFFENSIVE_WIN_L4_CSUM_DIFF',
       'TACKLES_L4_CSUM_DIFF', 'TACKLES_INSIDE50_L4_CSUM_DIFF',
       'ONE_PERCENTERS_L4_CSUM_DIFF', 'REBOUND_INSIDE50S_L4_CSUM_DIFF',
       'INTERCEPTS_L4_CSUM_DIFF', 'PRESSURE_L4_CSUM_DIFF',
       'PRESSURE_DEFENSEHALF_L4_CSUM_DIFF', 'SPOILS_L4_CSUM_DIFF',
       'DISPOSALS_EFFICIENCY_L4_CSUM_DIFF', 'KICKS_EFFICIENCY_L4_CSUM_DIFF',
       'KICK2HANDBALL_L4_CSUM_DIFF', 'POSESSIONS_CONTESTED_RATE_L4_CSUM_DIFF',
       'GOALS_ACCURACY_L4_CSUM_DIFF', 'HITOUTS_ADVANTAGE_RATE_L4_CSUM_DIFF',
       'CONTEST_OFFENSIVE_WIN_RATE_L4_CSUM_DIFF',
       'CONTEST_DEFENSIVE_LOSS_RATE_L4_CSUM_DIFF', 'AT_HOME', 'HOMEGROUND',
       'INTERSTATE')
```

```{r}
# RUN ONLY FOR SINGLE MODELS

afl <- data_build(homepath, awaypath, fixedpath)
resu <- fit_BTM_TSTV(afl, feats_L4)
resu$summary
```


```{r}
# RUN FOR 2-WINDOW

homepath15 <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2022.xlsx"
awaypath15 <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2022.xlsx"
homepath16 <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2023.xlsx"
awaypath16 <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2023.xlsx"
afl2015 <- data_build(homepath15, awaypath15, fixedpath)
afl2016 <- data_build(homepath16, awaypath16, fixedpath)

afl2015$home <- do.call(rbind, list(afl2015$home,
                                    afl2016$home))
afl2015$away <- do.call(rbind, list(afl2015$away,
                                    afl2016$away))
resu <- fit_BTM_TSTV(afl2015, feats_L4)
resu$summary
```

# PREDICTING WITH TEAM-SPECIFIC, TIME-VARIANT BT EXPANSION

predict_same_next_TSTV is used to predict with the team-specific, time-variant 
BT expansion on the train and test data.

Parameters:
  - model_TSTV (model): final model with significant features.
  - same_data (list): training data.
  - next_data (list): test data.

Returns:
  - accuracy_now (numeric): classification accuracy over train set
  - accuracy_next (numeric): classification accuracy over test set
  - conf_matrix_now: confusion matrix for the predictions over the train set.
  - conf_matrix_next: confusion matrix for the predictions over the test set.

```{r}
predict_same_next_TSTV <- function(model_TSTV, same_data,next_data) {
  
  pred_now_p <- predict(model_TSTV, se.fit = TRUE, level = 0, 
                      type = "response", newdata = same_data)$fit
  pred_now_b <- ifelse(pred_now_p > 0.5, 1, 0)
  pred_next_p <- predict(model_TSTV, se.fit = TRUE, level = 0, 
                      type = "response", newdata = next_data)$fit
  pred_next_b <- ifelse(pred_next_p > 0.5, 1, 0)
  accuracy_now <- mean(pred_now_b == same_data$HOME_WIN)
  accuracy_next <- mean(pred_next_b == next_data$HOME_WIN)
  conf_matrix_now <- confusionMatrix(data = factor(pred_now_b,levels = c(0, 1)), 
                               reference = factor(same_data$home$WIN,levels = c(0, 1)),
                               positive = '1')
  conf_matrix_next <- confusionMatrix(data = factor(pred_next_b,levels = c(0, 1)), 
                             reference = factor(next_data$home$WIN,levels = c(0, 1)),
                             positive = '1')
  
  result <- list(
    accuracy_now = accuracy_now,
    accuracy_next = accuracy_next,
    conf_matrix_now = conf_matrix_now,
    conf_matrix_next = conf_matrix_next
  )
  return(result)
  
}

```

# EXAMPLE

```{r}
test_homepath <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2023.xlsx"
test_awaypath <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2023.xlsx"
test <- data_build(test_homepath, test_awaypath, fixedpath)

prepre <- predict_same_next_TSTV(resu$model, afl2015, test)
prepre$conf_matrix_now
prepre$conf_matrix_next

```

# INTERACTION EFFECTS

Use these cells to test for interaction effects. The thesis found no significant
ones.

```{r}
homepath <- "E:/TFG DATA/TS_TV_DATA/HOME_ALL_AFL_TS_TV_2015.xlsx"
awaypath <- "E:/TFG DATA/TS_TV_DATA/AWAY_ALL_AFL_TS_TV_2015.xlsx"
afl <- data_build(homepath, awaypath, fixedpath)
resu <- fit_BTM_TSTV(afl, feats_ALL)
resu$summary
```

```{r}
mod_interactions <-  BTm(outcome = WIN,
                     player1 = home,
                     player2 = away,
                     formula = ~ AT_HOME
                     + HOMEGROUND
                     + HOMEGROUND:AT_HOME
                     # + INTERSTATE
                     # + AT_HOME:INTERSTATE
                     + (1|TEAM),
                     id = "TEAM",
                     data = afl)
summary(mod_interactions)
```

```{r}
resu$model <- BTm(outcome = WIN, player1 = home, player2 = away,
                      formula = ~ LADDERLY_POSITION_DIFF
                      + POINTSFOR_DIFF 
                      + HITOUTS_WIN_RATE_L4_CSUM_DIFF
                      + INTERSTATE
                      + AT_HOME
                      + HOMEGROUND
                      + AT_HOME:HOMEGROUND
                        + (1|TEAM),
                     id = "TEAM",
                     data = afl)
  
summary(resu$model)
```

```{r}
test_homepath <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2016.xlsx"
test_awaypath <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2016.xlsx"
test <- data_build(test_homepath, test_awaypath, fixedpath)

prepre <- predict_same_next_TSTV(resu$model, afl, test)
prepre$conf_matrix_now
prepre$conf_matrix_next
```

# GAME2GAME PREDICTION

dynamic_tester can be used to predict outcomes game-to-game. The thesis 
discourages this option since some games take place at the same time. Please, 
refer to dynamic_tester_ROUND

```{r}
dynamic_tester <- function(model, train_data, next_season, fixed, feats) {
  
  home_win_target <- next_season$home$WIN
  df_for_cm <- data.frame(TARGET = next_season$home$WIN, 
                           ROUND_NUMBER = next_season$home$ROUND_NUMBER)
  
  print(length(df_for_cm$TARGET))
  current_game <- 1
  num_games <- nrow(next_season$home)
  # print(num_games)
  list_preds <- vector("list", length = num_games)
  cat('GAMES TO PREDICT',num_games, "\n")
  
  while (current_game < num_games + 1) {
    
    # PREDICT NEXT GAME
    home_df <- next_season$home[current_game, ]
    away_df <- next_season$away[current_game, ]
    tester <- list(home = home_df, away = away_df, fixed = fixed)
    
    pred <- predict(model, se.fit = TRUE, level = 0, type = "response", newdata = tester)$fit
    list_preds[[current_game]] <- pred
    cat('PREDICTED GAME',current_game,
        "WITH ESTIMATED WINNING CHANCE",pred,
        "CURRENT LENGTH",length(list_preds),"\n")
    # RETRAIN MODEL WITH THIS GAME
    train_data$home <- do.call(rbind, list(train_data$home, home_df))
    train_data$away <- do.call(rbind, list(train_data$away, away_df))
    model <- fit_BTM_TSTV(train_data, feats)$model
    
    current_game <- current_game + 1

    
  }
  
  # RETURN PREDICTIONS
  print('---------------------------------')
  print('TESTING FINISHED')
  print('---------------------------------')
  list_preds <- lapply(list_preds, function(x) ifelse(x > 0.5, 1, 0))
  df_for_cm$preds <- unlist(list_preds)
  df_for_cm$correct <- ifelse(df_for_cm$TARGET == df_for_cm$preds, 1, 0)
  cm <- confusionMatrix(data = factor(df_for_cm$preds,levels = c(0, 1)), 
                        reference = factor(df_for_cm$TARGET,levels = c(0, 1)),
                        positive = '1')
  result <- list(df = df_for_cm,
                 cm = cm)
  
  return(result)
  
}
```



```{r}
fixedpath <- "E:/TFG DATA/last_ladder_position.xlsx"

homepath <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2016.xlsx"
awaypath <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2016.xlsx"
train_data <- data_build(homepath, awaypath, fixedpath)
resu <- fit_BTM_TSTV(train_data, feats_L4)

homepath_test <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2017.xlsx"
awaypath_test <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2017.xlsx"
next_season <- data_build(homepath_test, awaypath_test, fixedpath)

resu <- fit_BTM_TSTV(train_data, feats_L4)
```

```{r}
new_accuracy <- dynamic_tester(resu$model,
                               train_data, 
                               next_season, 
                               fixedpath,
                               feats_L4)
```
```{r}
new_accuracy
```


```{r}
write.xlsx(resu_cs$AIC_dataframe, 
           file = "E:/TFG DATA/RESULTS/AIC_PREDS_CS_w4.xlsx", row.names = FALSE)
```


# ROUND2ROUND PREDICTION

dynamic_tester_ROUND is used to predict games round-by-round with the Addition
or Subtitution strategies. Please, refer to the thesis for an explanation of 
these strategies.

Parameters:

  - model (model): team-specific, time-variant BT expansion trained beforehand.
  - train_data (list): data used for training the model before.
  - next_season (list): test data.
  - fixed (string): path to fixed effects data.
  - feats (vector): list of features to fit to the model.
  - subti (numeric): 1 applies the substitution strategy, otherwise applies the
  addition strategy.
  
Details:
  - Predicts on the test dataset round-by-round.
  - Adds or substitutes that round depending on the strategy specified in subti.
  - Retrains the model and predicts the following round.

Returns:
  - df (data.frame): dataframe with target variable, round number, prediction
   and whether the prediction was correct
  - cm: confusion matrix for predicted values.

```{r}
dynamic_tester_ROUND <- function(model, train_data, next_season, fixed, feats,
                                 subti) {
  
  home_win_target <- next_season$home$WIN
  df_for_cm <- data.frame(TARGET = next_season$home$WIN, 
                          ROUND_NUMBER = next_season$home$ROUND_NUMBER)
  
  current_round <- 1
  num_rounds <- max(next_season$home$ROUND_NUMBER)
  list_preds <- vector("list", length = nrow(next_season$home))
  cat('ROUNDS TO PREDICT', num_rounds, "\n",
      'GAMES TO PREDICT',length(df_for_cm$TARGET) , "\n")
  
  while (current_round <= num_rounds) {
    
    # PREDICT NEXT ROUND
    home_df <- next_season$home[next_season$home$ROUND_NUMBER == current_round, ]
    away_df <- next_season$away[next_season$away$ROUND_NUMBER == current_round, ]
    tester <- list(home = home_df, away = away_df, fixed = fixed)
    
    pred <- predict(model, se.fit = TRUE, level = 0, 
                    type = "response", newdata = tester)$fit
    list_preds[[current_round]] <- pred
    cat('PREDICTED ROUND', current_round,
        "WITH ESTIMATED WINNING CHANCE", pred,"\n")
    
    # RETRAIN MODEL WITH THIS ROUND
    
    ######################################################################
    if (subti == 1){
      train_data$home <- subset(train_data$home, ROUND_NUMBER != current_round)
      train_data$away <- subset(train_data$away, ROUND_NUMBER != current_round)
    }
    ######################################################################
    
    train_data$home <- do.call(rbind, list(train_data$home, home_df))
    train_data$away <- do.call(rbind, list(train_data$away, away_df))
    model <- fit_BTM_TSTV(train_data, feats)$model
    
    current_round <- current_round + 1
  }
  
  # RETURN PREDICTIONS
  cat('---------------------------------', '\n',
      'TESTING FINISHED', '\n',
      '---------------------------------')
  df_for_cm$probs_og <- unlist(list_preds)
  list_preds <- lapply(list_preds, function(x) ifelse(x > 0.5, 1, 0))
  df_for_cm$preds <- unlist(list_preds)
  df_for_cm$correct <- ifelse(df_for_cm$TARGET == df_for_cm$preds, 1, 0)
  cm <- confusionMatrix(data = factor(df_for_cm$preds, levels = c(0, 1)), 
                        reference = factor(df_for_cm$TARGET, levels = c(0, 1)),
                        positive = '1')
  result <- list(df = df_for_cm,
                 cm = cm)
  
  return(result)
}

```

# EXAMPLE

```{r}
fixedpath <- "E:/TFG DATA/last_ladder_position.xlsx"

homepath1 <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2022.xlsx"
awaypath1 <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2022.xlsx"
homepath2 <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2023.xlsx"
awaypath2 <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2023.xlsx"
afl1 <- data_build(homepath1, awaypath1, fixedpath)
afl2 <- data_build(homepath2, awaypath2, fixedpath)

afl1$home <- do.call(rbind, list(afl1$home,
                                    afl2$home))
afl1$away <- do.call(rbind, list(afl1$away,
                                    afl2$away))


homepath <- "E:/TFG DATA/TS_TV_DATA/HOME_ALL_AFL_TS_TV_2015.xlsx"
awaypath <- "E:/TFG DATA/TS_TV_DATA/AWAY_ALL_AFL_TS_TV_2015.xlsx"
train_data <- data_build(homepath, awaypath, fixedpath)
resu <- fit_BTM_TSTV(train_data, feats_ALL)

homepath_test <- "E:/TFG DATA/TS_TV_DATA/HOME_ALL_AFL_TS_TV_2016.xlsx"
awaypath_test <- "E:/TFG DATA/TS_TV_DATA/AWAY_ALL_AFL_TS_TV_2016.xlsx"
next_season <- data_build(homepath_test, awaypath_test, fixedpath)

new_accuracy <- dynamic_tester_ROUND(resu$model,
                               train_data, 
                               next_season, 
                               fixedpath,
                               feats_ALL)
new_accuracy$cm
```

```{r}
new_accuracy$cm
```


```{r}
write.xlsx(new_accuracy$df, 
           file = "E:/TFG DATA/RESULTS/EXPERIMENT5/ALL_ROUND_BY_ROUND_PREDS2.xlsx",
           row.names = FALSE)

```

# MIXED PREDICTION

dynamic_tester_MIXED is used to predict seasons with the Incremental strategy.
Please, refer to the thesis for an explanation of 
this strategy.

Parameters:

  - model (model): team-specific, time-variant BT expansion trained beforehand.
  - train_data (list): data used for training the model before.
  - next_season (list): test data.
  - fixed (string): path to fixed effects data.
  - feats (vector): list of features to fit to the model.
  - subti (numeric): 1 applies the substitution strategy, otherwise applies the
  incremental strategy.

Details:
  - Predicts the first 2 rounds with the previously trained model.
  - For round 3 onwards, it trains a model in the test season. If no features
  are found to be significant, it uses the model before. Otherwise, it predicts
  with the new model.
  
Returns:
  - df (data.frame): dataframe with target variable, round number, predictions
    for both models and whether they were correct

```{r}
dynamic_tester_MIXED <- function(model,
                                 train_data,
                                 next_season,
                                 fixed,
                                 feats,
                                 subti) {
  
  home_win_target <- next_season$home$WIN
  df_for_cm <- data.frame(TARGET = next_season$home$WIN, 
                          ROUND_NUMBER = next_season$home$ROUND_NUMBER)
  
  current_round <- 1
  num_rounds <- max(next_season$home$ROUND_NUMBER)
  list_preds <- vector("list", length = nrow(next_season$home))
  
  next_season_train <- list()
  list_preds_next_season <- vector("list", length = nrow(next_season$home))
  
  cat('ROUNDS TO PREDICT', num_rounds, "\n",
      'GAMES TO PREDICT',length(df_for_cm$TARGET) , "\n")
  
  while (current_round <= num_rounds) {
    
    # PREDICT NEXT ROUND
    home_df <- next_season$home[next_season$home$ROUND_NUMBER == current_round, ]
    away_df <- next_season$away[next_season$away$ROUND_NUMBER == current_round, ]
    tester <- list(home = home_df, away = away_df, fixed = fixed)
    
    pred <- predict(model, se.fit = TRUE, level = 0, 
                    type = "response", newdata = tester)$fit
    list_preds[[current_round]] <- pred
    
    if (current_round == 1){
      pred_same <- pred
      list_preds_next_season[[current_round]] <- pred_same
      next_season_train <- list(home = next_season$home[next_season$home$ROUND_NUMBER == 1, ],
                                 away = next_season$away[next_season$away$ROUND_NUMBER == 1, ],
                                 fixed = fixed)
    } else if (current_round == 2){
      pred_same <- pred
      list_preds_next_season[[current_round]] <- pred_same
      next_season_train$home <- do.call(rbind, list(next_season_train$home, home_df))
      next_season_train$away <- do.call(rbind, list(next_season_train$away, away_df))
    }
    else {
      if (length(model_same) == 0){
        cat('Model for round', current_round,
            'has no significant features. Using historical model')
        pred_same <- pred
        list_preds_next_season[[current_round]] <- pred_same
        next_season_train$home <- do.call(rbind, list(next_season_train$home, home_df))
        next_season_train$away <- do.call(rbind, list(next_season_train$away, away_df))
        
      } else{
        pred_same <- predict(model_same, se.fit = TRUE, level = 0, type = "response", newdata = tester)$fit
        list_preds_next_season[[current_round]] <- pred_same
        next_season_train$home <- do.call(rbind, list(next_season_train$home, home_df))
        next_season_train$away <- do.call(rbind, list(next_season_train$away, away_df))
      }
      
    }
    
    cat('PREDICTED ROUND', current_round,
        "WITH ESTIMATED WINNING CHANCE", pred,
        'AND', pred_same,"\n")
    
    # RETRAIN MODEL WITH THIS ROUND
    
    ######################################################################
    if (subti == 1){
      train_data$home <- subset(train_data$home, ROUND_NUMBER != current_round)
      train_data$away <- subset(train_data$away, ROUND_NUMBER != current_round)
    }
    # Comment for round addition, uncomment for round substitution
    # train_data$home <- subset(train_data$home, ROUND_NUMBER != current_round)
    # train_data$away <- subset(train_data$away, ROUND_NUMBER != current_round)
    ######################################################################
    
    train_data$home <- do.call(rbind, list(train_data$home, home_df))
    train_data$away <- do.call(rbind, list(train_data$away, away_df))
    model <- fit_BTM_TSTV(train_data, feats)$model
    
    if (current_round != 1){
      model_same <- fit_BTM_TSTV(next_season_train, feats)$model
    }
    
    current_round <- current_round + 1
  }
  
  
  # RETURN PREDICTIONS
  cat('---------------------------------', '\n',
      'TESTING FINISHED', '\n',
      '---------------------------------')
  df_for_cm$probs_og <- unlist(list_preds_next_season)
  list_preds <- lapply(list_preds, function(x) ifelse(x > 0.5, 1, 0))
  list_preds_same <- lapply(list_preds_next_season, function(x) ifelse(x > 0.5, 1, 0))
  df_for_cm$preds <- unlist(list_preds)
  df_for_cm$preds_same <- unlist(list_preds_same)
  df_for_cm$coincide <- ifelse(df_for_cm$preds == df_for_cm$preds_same, 1, 0)
  df_for_cm$correct <- ifelse(df_for_cm$TARGET == df_for_cm$preds, 1, 0)
  df_for_cm$correct_same <- ifelse(df_for_cm$TARGET == df_for_cm$preds_same, 1, 0)
  # cm <- confusionMatrix(data = factor(df_for_cm$preds, levels = c(0, 1)), 
  #                       reference = factor(df_for_cm$TARGET, levels = c(0, 1)),
  #                       positive = '1')
  result <- list(df = df_for_cm
                 # cm = cm
                 )
  
  return(result)
}
```

```{r}
fixedpath <- "E:/TFG DATA/last_ladder_position.xlsx"

homepath <- "E:/TFG DATA/TS_TV_DATA/HOME_ALL_AFL_TS_TV_2015.xlsx"
awaypath <- "E:/TFG DATA/TS_TV_DATA/AWAY_ALL_AFL_TS_TV_2015.xlsx"
train_data <- data_build(homepath, awaypath, fixedpath)
resu <- fit_BTM_TSTV(train_data, feats_ALL)

homepath_test <- "E:/TFG DATA/TS_TV_DATA/HOME_ALL_AFL_TS_TV_2016.xlsx"
awaypath_test <- "E:/TFG DATA/TS_TV_DATA/AWAY_ALL_AFL_TS_TV_2016.xlsx"
next_season <- data_build(homepath_test, awaypath_test, fixedpath)

new_accuracy <- dynamic_tester_MIXED(resu$model,
                               train_data, 
                               next_season, 
                               fixedpath,
                               feats_ALL)
```

```{r}
write.xlsx(new_accuracy$df, 
           file = "E:/TFG DATA/RESULTS/EXPERIMENT5/ALLMIXED15_16.xlsx",
           row.names = FALSE)
```
# MAJORITY VOTING

The Majority Voting predictions are done through an orchestrator instead of a
function. The predictions of different models or strategies are combined: if 3
of them predict the HOME_WIN probability to be above 0.5, the prediction of the 
Majority Voting is is 1. Otherwise it is 0.

The cells below orchestrate the Majority Voting strategy and create a data frame
to compare the predictions and accuracy by round number and team.

```{r}
# This is not a function, but an orchestrator

###################################################################
# Train and test contest-specific model
###################################################################
train_cs <- read.xlsx("E:/TFG DATA/ST_DATA/AFL_ST_2015.xlsx")
test_cs <- read.xlsx("E:/TFG DATA/ST_DATA_TEST/AFL_ST_2016_TEST.xlsx")

train_cs$HOME_TEAM <- as.factor(train_cs$HOME_TEAM)
train_cs$AWAY_TEAM <- as.factor(train_cs$AWAY_TEAM)
train_cs$HOME_WIN <- as.numeric(as.character(train_cs$HOME_WIN))
train_cs$AWAY_WIN <- as.numeric(as.character(train_cs$AWAY_WIN))
train_cs$HOME_TEAM <- data.frame(team = train_cs$HOME_TEAM,
                                     AT_HOME = 1)
train_cs$AWAY_TEAM <- data.frame(team = train_cs$AWAY_TEAM
                                      , AT_HOME = 0)

test_cs$HOME_TEAM <- as.factor(test_cs$HOME_TEAM)
test_cs$AWAY_TEAM <- as.factor(test_cs$AWAY_TEAM)
test_cs$HOME_WIN <- as.numeric(as.character(test_cs$HOME_WIN))
test_cs$AWAY_WIN <- as.numeric(as.character(test_cs$AWAY_WIN))
test_cs$HOME_TEAM <- data.frame(team = test_cs$HOME_TEAM,
                                     AT_HOME = 1)
test_cs$AWAY_TEAM <- data.frame(team = test_cs$AWAY_TEAM
                                      , AT_HOME = 0)


reference_w1 <- c("GEEL", "PORT", "WB",
               "PORT", "PORT","GWS",
               "STK","STK","ADEL",
               "ADEL")


model_CS <- BTm(outcome = cbind(HOME_WIN, AWAY_WIN), 
                          player1 = HOME_TEAM, player2 = AWAY_TEAM, 
                          formula = ~team + AT_HOME, id = "team",
                           refcat = reference_w1[1],
                      data = train_cs)

pred_CS <- predict(model_CS, se.fit = TRUE, level = 0, 
                      type = "response", newdata = test_cs)$fit


###################################################################
# Train and test complex BT FULL YEAR
###################################################################


homepath1 <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2015.xlsx"
awaypath1 <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2015.xlsx"
homepath2 <- "E:/TFG DATA/TS_TV_DATA/HOME_L4_AFL_TS_TV_2016.xlsx"
awaypath2 <- "E:/TFG DATA/TS_TV_DATA/AWAY_L4_AFL_TS_TV_2016.xlsx"
fixedpath <- "E:/TFG DATA/last_ladder_position.xlsx"
train_TVTS <- data_build(homepath1, awaypath1, fixedpath)
test_TVCS <- data_build(homepath2, awaypath2, fixedpath)

model_TV_TS <- fit_BTM_TSTV(train_TVTS, feats_L4)$model
pred_TVTS_full_season <- predict(model_TV_TS, se.fit = TRUE, level = 0, 
                      type = "response", newdata = test_TVCS)$fit

################################
# ADDITION

preds_addition <- dynamic_tester_ROUND(model_TV_TS,train_TVTS, test_TVCS,
                                  fixedpath, feats_L4, subti = 0)$df$probs_og
# SUBTITUTION
preds_substitution <- dynamic_tester_ROUND(model_TV_TS,train_TVTS, test_TVCS,
                                  fixedpath, feats_L4, subti = 1)$df$probs_og

# INCREMENTAL
preds_incremental <- dynamic_tester_MIXED(model_TV_TS,train_TVTS, test_TVCS,
                                  fixedpath, feats_L4, subti = 0)$df$probs_og


```

```{r}
MV_df <-data.frame(ROUND_NUMBER = test_TVCS$home$ROUND_NUMBER,
                   HOME_TEAM = test_TVCS$home$TEAM,
                   AWAY_TEAM = test_TVCS$away$TEAM,
                    TARGET = test_cs$HOME_WIN,
                   CS_PREDICTION = pred_CS,
                   TV_TS_PREDICTION = pred_TVTS_full_season,
                   ADDITION_PREDICTION = preds_addition,
                   SUBSTITUTION_PREDICTION = preds_substitution,
                   INCREMENTAL_PREDICTION = preds_incremental)

MV_df$MAJORITY_VOTE <- ifelse(rowSums(MV_df[,
                            c("CS_PREDICTION", "TV_TS_PREDICTION",
                              "ADDITION_PREDICTION", "SUBSTITUTION_PREDICTION", 
                              "INCREMENTAL_PREDICTION")
                            ] > 0.5) >= 3, 1, 0)

MV_df$CS_correct <- ifelse(MV_df$TARGET == ifelse(MV_df$CS_PREDICTION > 0.5, 1, 0), 1, 0)
MV_df$TVTS_correct <- ifelse(MV_df$TARGET == ifelse(MV_df$TV_TS_PREDICTION > 0.5, 1, 0), 1, 0)
MV_df$ADD_correct <- ifelse(MV_df$TARGET == ifelse(MV_df$ADDITION_PREDICTION > 0.5, 1, 0), 1, 0)
MV_df$SUBS_correct <- ifelse(MV_df$TARGET == ifelse(MV_df$SUBSTITUTION_PREDICTION > 0.5, 1, 0), 1, 0)
MV_df$INCR_correct <- ifelse(MV_df$TARGET == ifelse(MV_df$INCREMENTAL_PREDICTION > 0.5, 1, 0), 1, 0)
MV_df$MV_correct <- ifelse(MV_df$TARGET == MV_df$MAJORITY_VOTE, 1, 0)

MV_df$num_VOTED_YES <- rowSums(MV_df[,
                            c("CS_PREDICTION", "TV_TS_PREDICTION",
                              "ADDITION_PREDICTION", "SUBSTITUTION_PREDICTION", 
                              "INCREMENTAL_PREDICTION")
                            ] > 0.5)

CONF_MATRIX <- confusionMatrix(data = factor(MV_df$MAJORITY_VOTE,levels = c(0, 1)), 
                               reference = factor(MV_df$TARGET,levels = c(0, 1)),
                               positive = '1')
CONF_MATRIX
```

```{r}

write.xlsx(MV_df, 
           file = "E:/TFG DATA/RESULTS/EXPERIMENT5/MV15_16.xlsx",
           row.names = FALSE)
```

